\chapter{Introduction}\label{introduction}

\begin{quote} 
  \emph{The aim of this work is to build a Blockchain based system to manage
  the identity of patients and find out the suitability of creating such a
  system in the Healthcare environment.  Traditional databases alone can become
  vulnerable and a target to groups of malicious actors. Also the data that
  forms the identity of a pattient is often fragmented across multiple
  Healthcare organizations, in such a way that, to a get an oveview of the
  patients history and diagnosis you would need to merge together all the
  pieces of information stored in these data systems. Blockchain has a set of
  features could improve upon these aspects if used together with existing
  systems. This thesis provides an insight into the design and implementation
  of a Blockchain based system for managing the identity of patients in a
  Healthcare context and its subsequent evaluation.  The system should allow a
  patient to be able to manage his data and control its access.  It can be used
  to handle the patient’s medical identity, for example, in hospitals or
  clinics alongside the current Information Systems available in nowadays
  regular medical environments. The creation of this system and its subsequent
  evaluation will provide interesting conclusions to medical staff as well as
  patients, regarding its potential implementation and deployment in the
  field.} 
\end{quote}

Health is intrinsically linked with technology as new technologies enable safer
and better treatments. It is also worth noting that computing devices and
networks are now easily available and widespread to the general population in
more develop countries. Healthcare organizations now store patients data on a
digital format. The Electronic Health Record (\textbf{EHR}) is an abstract
concept representing the patients digitally stored clinical data and their
identity in a medical context.

Electronic Health Records (\textbf{EHR}) for many years have lacked a standard,
even tough, more recently there has been some progress regarding this matter.
Standards are an important step to implement because they allow
interoperability between different organizations. The Health Level 7
(\textbf{HL7}) standard, being developed by Health Level Seven organization, is
growing in use and is represented internationally. Health Level Seven Portugal
is starting its operations and is building a community to support the use of
this standard in Portugal~\cite{HealthLevel7}.

Blockchain is often known as the technology behind the Bitcoin Cryptocurrency.
The phenomenon we know as Bitcoin depends on two complementary technologies,
digital tokens and Blockchain, that together facilitate trust, immutability and
resiliency~\cite{Evans2016}.

A Blockchain runs on a network of computers and has a list of records that are
replicated across the participating peers. Blockchain, as we know today, was
conceptualized as the public ledger~\footnote{A ledger is defined as an object
in which items are regularly recorded, originally business activities and money
received or paid, but in reality, it can be used to store any type of record.}
for the Bitcoin cryptocurrency in 2008 by Satoshi Nakamoto, a pen name of, a
still unknown to this day, individual or organization of individuals. The
network was implemented in 2009 and many are now finding it has a much broader
potential across many fields, with some implementations even resembling a
programming platform to execute code in an autonomous
manner~\cite{Nakamoto2008}.

Traditional databases and architectures can become vulnerable and a target to
groups of malicious actors that possess the technical expertise to deny
services with Distributed Denial of Service (\textbf{DDOS}) attacks~\footnote{A
Distributed Denial of Service attack is an attempt to make an online service
unavailable by overwhelming it with traffic from multiple sources.} or cause a
data breach~\footnote{A data breach is the intentional or unintentional release
of secure or private/confidential information to an untrusted environment.}. 

Making matters worse other problems spring to mind. The data that forms the
identity of a patient is often fragmented across multiple Healthcare
organizations, in such a way that, to get a true overview of the patients
history and diagnosis you would need to merge together all the pieces of
information stored in data systems that are hosted in architecturally different
and sometimes competing Healthcare Information Systems. Transparency is also a
concern, as a patient does not currently possess the means to track how his
medical data is being handled by the medical services he used.

As more information becomes available, new insights can be extracted by health
professionals that lead to an overall improvement of the patients interaction
with the Healthcare ecosystem. However, maintaining this huge amount of data
secure is a costly and risky matter for every party involved. Security and
privacy should be a top concern regarding this sensitive data. 

In this document different Blockchain implementations are explored to get an
overview of their capabilities and purposes. Also, some use cases of Blockchain
based systems used in the Healthcare field are presented. More precisely, in
Chapter~\ref{background}, a brief introduction to Blockchain is made followed
by an introduction to its most prominent implementations. The technology is
further explored in Chapter~\ref{blockchain} and a number of real-world use
cases of this technology in the Healthcare field are explored.  In
Chapter~\ref{development} a prototype system to evaluate the usability of this
technology in the Healthcare field is design, implemented and built as well as
evaluated. Finally, in Chapter~\ref{Conclusion}, some conclusions are observed
regarding the change enabled by these advances.
